Solving the maximum independant set problem with einsum networks

### 2021-06-02 paper/paper.tex
update example

\subsection{Enumerating configurations}
\begin{lstlisting}
using OMEinsum, TropicalNumbers

struct ConfigEnumerator{N}
    data::Vector{BitVector}
end

Base.length(x::ConfigEnumerator{N}) where N = length(x.data)
Base.:(==)(x::ConfigEnumerator, y::ConfigEnumerator) = x.data == y.data

function Base.:+(x::ConfigEnumerator{N}, y::ConfigEnumerator{N}) where N
    res = ConfigEnumerator{N}(vcat(x.data, y.data))
    @show length(x), length(y), length(res)
    @assert length(res) == length(x) + length(y)
    return res
end

function Base.:*(x::ConfigEnumerator{L}, y::ConfigEnumerator{L}) where L
    M, N = length(x), length(y)
    z = Vector{BitVector}(undef, M*N)
    for j=1:N, i=1:M
        z[(j-1)*M+i] = x.data[i] .| y.data[j]
    end
    #@show M, N, length(z)
    @assert length(z) == length(x) * length(y)
    return ConfigEnumerator{L}(z)
end

Base.zero(::Type{ConfigEnumerator{N}}) where N = ConfigEnumerator{N}(BitVector[])
Base.one(::Type{ConfigEnumerator{N}}) where N = ConfigEnumerator{N}([falses(N)])

function onehot(::Type{ConfigEnumerator{N}}, i::Int) where N
    res = one(ConfigEnumerator{N})
    res.data[1][i] |= true
    return res
end

function enumerator_t(::Type{T}, ix::NTuple{2}, vertex_index) where {T1, N, T<:CountingTropical{T1, ConfigEnumerator{N}}}
    [one(T) one(T); one(T) zero(T)]
end
function enumerator_t(::Type{T}, ix::NTuple{1}, vertex_index) where {T1, N, T<:CountingTropical{T1, ConfigEnumerator{N}}}
    [one(T), CountingTropical(one(T1), onehot(ConfigEnumerator{N}, vertex_index[ix[1]]))]
end

symbols(::EinCode{ixs}) where ixs = unique(Iterators.flatten(ixs))
symbols(ne::NestedEinsum) = symbols(Iterators.flatten(ne))
function mis_enumerate(code)
    vertex_index = Dict([s=>i for (i, s) in enumerate(symbols(code))])
	xs = generate_xs!((T, ix)->enumerator_t(T, ix, vertex_index), CountingTropical{Float64,ConfigEnumerator{length(vertex_index)}}, code, Vector{Any}(undef, ninput(code)))
	code(xs...)
end
\end{lstlisting}
\end{document}

>>>>

\subsection{Enumerating configurations}

The following is a complete example.

\begin{lstlisting}
using OMEinsum, TropicalNumbers
using OMEinsum: NestedEinsum

struct ConfigEnumerator{N,C}
    data::Vector{StaticBitVector{N,C}}
end

Base.length(x::ConfigEnumerator{N}) where N = length(x.data)
Base.:(==)(x::ConfigEnumerator{N,C}, y::ConfigEnumerator{N,C}) where {N,C} = x.data == y.data

function Base.:+(x::ConfigEnumerator{N,C}, y::ConfigEnumerator{N,C}) where {N,C}
    res = ConfigEnumerator{N,C}(vcat(x.data, y.data))
    return res
end

function Base.:*(x::ConfigEnumerator{L,C}, y::ConfigEnumerator{L,C}) where {L,C}
    M, N = length(x), length(y)
    z = Vector{StaticBitVector{L,C}}(undef, M*N)
    for j=1:N, i=1:M
        z[(j-1)*M+i] = x.data[i] .| y.data[j]
    end
    return ConfigEnumerator{L,C}(z)
end

Base.zero(::Type{ConfigEnumerator{N,C}}) where {N,C} = ConfigEnumerator{N,C}(StaticBitVector{N,C}[])
Base.one(::Type{ConfigEnumerator{N,C}}) where {N,C} = ConfigEnumerator{N,C}([TropicalNumbers.staticfalses(StaticBitVector{N,C})])

function enumerator_t(::Type{T}, ix::NTuple{2}, vertex_index) where {T1, N, C, T<:CountingTropical{T1,ConfigEnumerator{N,C}}}
    [one(T) one(T); one(T) zero(T)]
end
function enumerator_t(::Type{T}, ix::NTuple{1}, vertex_index) where {T1, N, C, T<:CountingTropical{T1,ConfigEnumerator{N,C}}}
    [one(T), CountingTropical(one(T1), ConfigEnumerator([TropicalNumbers.onehot(StaticBitVector{N,C}, vertex_index[ix[1]])]))]
end
function enumerator_t(::Type{T}, ix::NTuple{2}, vertex_index) where {T1,N,C, T<:ConfigTropical{T1,N,C}}
    [one(T) one(T); one(T) zero(T)]
end
function enumerator_t(::Type{T}, ix::NTuple{1}, vertex_index) where {T1,N,C, T<:ConfigTropical{T1,N,C}}
    [one(T), T(one(T1), TropicalNumbers.onehot(StaticBitVector{N,C}, vertex_index[ix[1]]))]
end

function generate_xs!(f, x::T, code::NestedEinsum, xs) where {T}
    for (ix, arg) in zip(OMEinsum.getixs(code.eins), code.args)
		if arg isa Integer
			xs[arg] = f(x, ix)
		else
        	generate_xs!(f, x, arg, xs)
		end
    end
	return xs
end

function generate_xs!(f, x::T, code::EinCode, xs) where {T}
    for (i,ix) in enumerate(OMEinsum.getixs(code))
		xs[i] = f(x, ix)
    end
	return xs
end

symbols(::EinCode{ixs}) where ixs = unique(Iterators.flatten(ixs))
symbols(ne::NestedEinsum) = symbols(Iterators.flatten(ne))
ninput(::EinCode{ixs}) where ixs = length(ixs)
ninput(ne::NestedEinsum) = ninput(Iterators.flatten(ne))
function mis_config(code; all=false)
    vertex_index = Dict([s=>i for (i, s) in enumerate(symbols(code))])
    N = length(vertex_index)
    C = TropicalNumbers._nints(N)
    T = all ? CountingTropical{Float64, ConfigEnumerator{N,C}} : ConfigTropical{Float64, N, C}
	xs = generate_xs!((T, ix)->enumerator_t(T, ix, vertex_index), T, code, Vector{Any}(undef, ninput(code)))
	code(xs...)
end

mis_config(optimize_greedy(ein"i,j,k,l,ij,jk,kl,li->", Dict('i'=>2, 'j'=>2, 'k'=>2, 'l'=>2)); all=true)  # square
\end{lstlisting}

You will see the following output

\begin{lstlisting}
0-dimensional Array{CountingTropical{Float64, ConfigEnumerator{4, 1}}, 0}:
(2.0, ConfigEnumerator{4, 1}(StaticBitVector{4, 1}[1010, 0101]))ₜ
\end{lstlisting}

\end{document}
